Examining readability of text generated through GPT-3

### 2023-03-28 paper/paper.tex
play around
update paper

\begin{document}
\maketitle
\begin{abstract}
    This paper examines readability-related metrics of AI-generated text, specifically texts produced by OpenAI's GPT-3. Generation was performed across 10 topics and the 6 proficiency levels defined by the CEFR framework, and outputs were compared with Simple Wikipedia and (standard) Wikipedia articles on the same topics.
\end{abstract}

@@

It is however not obvious that AI generation tools reach the desired quality or adherence to the didactic goals that teachers set out. Owing to the increased popularity of these tools, more research is desirable on their applicability in a number of different settings.

This work aims to make a contribution to those questions by verifying the readability level of text content generated by GPT-3 with regard to 10 topics and the 6 proficiency levels defined by the CEFR framework. A number of complexity metrics are computed on the generated texts and compared with Simple Wikipedia\footnote{\url{https://simple.wikipedia.org/wiki/Main_Page}} as well as Wikipedia\footnote{\url{https://www.wikipedia.org/}} articles on the same topics in order to gauge the true readability levels of the model outputs.
\section{Dataset and metrics}
\subsection{Dataset}

@@

@@

>>>>

\begin{document}
\maketitle
\begin{abstract}
    This paper examines readability-related metrics of AI-generated text, specifically texts produced by OpenAI's GPT-3. Generation was performed for the English language across 10 topics and the 6 proficiency levels defined by the CEFR framework, and outputs were compared with Simple Wikipedia and (standard) Wikipedia articles on the same topics. 7 commonly used complexity metrics have been employed, targeting chiefly lexis and syntax. In the limited scope of this experiment, GPT-3 is observed to at least partially aware of what constitutes reading difficulty.
\end{abstract}

@@

It is however not obvious that AI generation tools reach the desired quality or adherence to the didactic goals that teachers set out. Owing to the increased popularity of these tools, more research is desirable on their applicability in a number of different settings. For example, GPT-3 has been observed to "make up" information about some real-world events or phenomena, a fact that makes it hard to rely on for teaching. In the specific context of this experiment, employing GPT-3 for language-learning material is only valuable insofar as the generated content is adequate in difficulty to the audience.

This work aims to make a contribution to these questions by verifying the readability level of text content generated by GPT-3 with regard to 10 topics and the 6 proficiency levels defined by the CEFR framework. A number of complexity metrics are computed on the generated texts and compared with Simple Wikipedia\footnote{\url{https://simple.wikipedia.org/wiki/Main_Page}} as well as Wikipedia\footnote{\url{https://www.wikipedia.org/}} articles on the same topics in order to gauge the true readability levels of the model outputs.
\section{Dataset and metrics}
\subsection{Dataset}

@@

Interesting discussions can be had around topics as well, since, depending on GPT-3's ability to generalize across domains, certain areas might be problem areas for it. For example, "semantics" and "quantum mechanics" would be hard to write about for A1 level students as they are likely to be children who have not interacted with those fields yet.

Generally, a difference in values is visible across the board, more so at the lower difficuly levels. At CEFR levels B2 and above, many of the metrics appear similar regardless of topic. Among the selected measures, NVR (noun to verb ratio) shows the highest cross-topic variation, maintaining strong differentitation even at the C2-level texts. Figure \ref{fig:wit-nvr-topics} shows metric values at the A1 and C2 levels, which indicate that topics commonly regarded as intrinsically have higher information density in GPT-3 texts. This is however a weak case, as few metrics differ so sharply as NVR, and bars in these figures correspond to a single text entry.

\input{figs/nvr_topics.tex}

@@

Since recent research (e.g. \citealp{Deutsch_2020}) has also focused on readability assessment through neural approaches without such metrics, this could also be explored with AI text data. An attempt was made in relationship to this experiment to train a readability analyzer with GLoVe word embeddings \citep{pennington2014glove} and Wikipedia data, but this unfortunaly was inconclusive, and thus results from this analyzer haven't been reported.

Mainly, however, future extensions of this experiment or related experiments would benefit from higher volume. Simply put, more data is necessary across multiple domains to make educated conclusions about the quality of GPT-3 outputs in a teaching setting. While data is easy to generate, quality annotations on it are yet unavailable. Indeed, it is a good question whether making any is a fruitful enterprise, as the actual models behind GPT-3 and ChatGPT are constantly updated - thus making it hard to pinpoint what exactly is being studied.
